### 2018-10-02 figures/01_county_extent.Rmd
add makefile figure infrastructure

---
title: "Figures"
output:
  pdf_document:
    fig_caption: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.path = "../figures/", fig.width = 5, fig.height = 6)
```

```{r read_explore_chunk, echo=FALSE, eval=TRUE}
knitr::read_chunk("../scripts/explore_lagos_ag.R")
```

```{r lagos_setup, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE}

```

```{r mendota_oneida, echo=FALSE, message=TRUE, eval=FALSE}
llids <- lake_info(lg, 
                   name = c("Lake Mendota", "Oneida Lake"), 
                   state = c("Wisconsin", "New York"))$lagoslakeid

cat(paste(c("Lake Mendota", "Oneida Lake")[which(llids %in% ep$lagoslakeid)], "is included in the hi ag sites."))
```

<!-- * Find counties that intersect large numbers of hi ag iws with tp + tn data -->

```{r county_extent, eval=TRUE, echo=FALSE, fig.cap = "Number of lake watersheds per county where lakes are limited to those with high agricultural land-use cover (> 40%) and water quality data in LAGOS measured at > 3 sampling events between 1995 and 2005.", warning=FALSE, fig.width = 5.5, fig.height = 3}

```