### 2018-10-03 figures/01_county_extent.Rmd
caption and number cleanup

---
output: pdf_document
header-includes: 
  - \pagenumbering{gobble}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.path = "../figures/", fig.width = 5, fig.height = 5)
```

```{r read_explore_chunk, echo=FALSE, eval=TRUE}
# setwd("figures")
knitr::read_chunk("../scripts/explore_lagos_ag.R")
```

```{r lagos_setup, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE}

```

```{r mendota_oneida, echo=FALSE, message=TRUE, eval=FALSE}
llids <- lake_info(lg, 
                   name = c("Lake Mendota", "Oneida Lake"), 
                   state = c("Wisconsin", "New York"))$lagoslakeid

cat(paste(c("Lake Mendota", "Oneida Lake")[which(llids %in% ep$lagoslakeid)], "is included in the hi ag sites."))
```

<!-- * Find counties that intersect large numbers of hi ag iws with tp + tn data -->

```{r 01_county_extent, eval=TRUE, echo=FALSE, fig.width = 5.5, fig.height = 3}

```