### 2018-10-11 figures/03_wetland_potential.Rmd
add 03_wetland_potential

---
output: pdf_document
header-includes: 
  - \pagenumbering{gobble}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.path = "../figures/", fig.width = 5, fig.height = 5)
```

```{r 03_wetland_potential, fig.height=3, echo=FALSE}
library(kableExtra)
library(dplyr)
library(tidyr)
library(magrittr)

library(LAGOSNE)
library(sf)
library(ggplot2)
library(cowplot)
library(concaveman)

states <- st_as_sf(maps::map("state", fill = TRUE, plot = FALSE))
ep      <- coordinatize(readRDS("../data/ep.rds"))
ep_bbox <- concaveman(ep)$polygons
states  <- states[unlist(lapply(st_intersects(states, ep_bbox), 
                               function(x) length(x) > 0)),]

# setwd("figures")
gssurgo_raw <- readRDS("../data/gssurgo/gssurgo.rds")$res
gssurgo_raw$llid <- as.integer(gssurgo_raw$llid)
lg <- lagosne_load("1.087.1")

gssurgo <- left_join(gssurgo_raw, 
                         dplyr::select(lg$locus, 
                         lagoslakeid, nhd_long, nhd_lat), 
                         by = c("llid" = "lagoslakeid"))
gssurgo <- left_join(gssurgo, 
            dplyr::select(lg$iws.conn, 
              wl_dissolved = iws_wl_allwetlandsdissolved_overlapping_area_pct, 
              wl_undissolved = iws_wl_allwetlandsundissolved_overlapping_area_pct,
              lagoslakeid), by = c("llid" = "lagoslakeid")) %>% 
          mutate(diff = wetland_potential - wl_dissolved) 
gssurgo <- coordinatize(gssurgo)

plot_grid(
  ggplot(data = gssurgo) +
    geom_point(aes(x = wetland_potential, 
                   y = wl_dissolved, color = diff > 30)) + 
    xlab("gssurgo wetland potential (%)") + ylab("lagos nwi wetland pct (%)") + 
    labs(color = "difference (%)") +
    geom_abline(slope = 1, intercept = 0) + 
    theme(legend.position = "none"),
  ggplot() + 
    geom_sf(data = states, fill = "white") +
    geom_sf(data = gssurgo, aes(color = diff > 30), size = 0.8) +
    theme_minimal() +
    theme(legend.position = "none"), labels = "AUTO"
)

```