### 2018-10-25 README.Rmd
rm old research qs

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "figures/",
  warning = FALSE,
  message = FALSE
)
```

Code and data for:

## The effect of crop type, fertilizer, and manure applications on lake nutrients in agriculture watersheds at regional scales

> Journal Target: Ecosystems

### Products

Figures: [manuscript/figures.pdf](manuscript/figures.pdf)

Tables: [manuscript/tables.pdf](manuscript/tables.pdf)

## Potential predictor variables

* LAGOS
    * Lake depth
    * IWS area
    * IWS/Lake area
* Cropland Data Layer
    * IWS pasture types (forage or not)
    * IWS specific crops (corn, wheat, etc.)
    * IWS crop frequency/rotation
* gSSURGO
    * IWS wetland potential (as a proxy for tile drainage)
    * IWS soil organic carbon
    * IWS root zone depth
    * IWS lithology (sand %, clay %, etc.)
* USGS
    * IWS from county - manure, fertilizer, and deposition of N,P

## Research questions

## Timeline

## Issues

* Availability of data products through time is highly variable. How to align or decide to treat properties as constant?

```{r maps, echo=FALSE}
knitr::include_graphics("figures/rmd-hi_ag_iws_w_ep-1.png")
```

```{r echo=FALSE}
cdl_key <- read.csv("data/cdl/cdl_key.csv", stringsAsFactors = FALSE)
```

```{r lagos_v_cdl_v_usgs, echo=FALSE, eval=FALSE}
# tp vs lagos ag, cdl ag, usgs P inputs

# ---- setup ----

library(dplyr)
library(ggplot2)
library(magrittr)
library(LAGOSNE)
library(cowplot)
library(sf)

lg          <- lagosne_load("1.087.1")
ep          <- readRDS("data/ep.rds")
cdl_summary <- read.csv("data/cdl/cdl_summary.csv", stringsAsFactors = FALSE)
usgs        <- readRDS("data/usgs/usgs.rds")

lg_ep <- mutate(lg$iws.lulc, 
                lg_ag = rowSums(cbind(iws_nlcd2006_pct_81, 
                                      iws_nlcd2006_pct_82), na.rm = TRUE)) %>%
  select(lagoslakeid, lg_ag) %>%
  left_join(ep, by = "lagoslakeid") %>%
  filter(!is.na(tp))

cdl_ep <- filter(cdl_summary, variable == "ag") %>%
  left_join(
    filter(ep, !is.na(tp)), by = c("llid" = "lagoslakeid")) %>%
  mutate(cdl_ag = value)

usgs_ep <- filter(usgs, !is.na(tp))

# ---- plotting ----

plot_grid(
ggplot() +
  geom_point(data = lg_ep, aes(x = lg_ag, y = tp)) +
  xlab("LAGOS (NLCD 2006) Ag (%)"),
ggplot() +
  geom_point(data = cdl_ep, aes(x = cdl_ag, y = tp)) +
  xlab("CDL Ag (%)"),
ggplot() + 
  geom_point(data = usgs_ep, aes(x = phosphorus_input, y = tp)) + 
  scale_x_log10() + 
  xlab("log(P Input)"))

```