### 2018-10-26 figures/04_nlcd-versus-cdl.Rmd
include pasture in cdl total ag

---
output: pdf_document
header-includes: 
  - \pagenumbering{gobble}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.path = "../figures/", fig.width = 5, fig.height = 5)
```

```{r 04_nlcd-versus-cdl, fig.height=3, echo=FALSE}
library(LAGOSNE)
library(ggplot2)
library(HydroData)
library(dplyr)
library(tidyr)
# setwd("data")
cdl <- read.csv("../data/cdl/cdl_summary.csv", stringsAsFactors = FALSE) %>%
  dplyr::select(lagoslakeid = llid, cdl_ag_99 = ag, cdl_pasture_81 = pasture)

lg  <- lagosne_load("1.087.1")$iws.lulc %>%
  dplyr::filter(lagoslakeid %in% cdl$lagoslakeid) %>%
  dplyr::select(lagoslakeid, matches(".*nlcd2011.*")) %>%
  dplyr::select(lagoslakeid, 
                nlcd_pasture_81 = iws_nlcd2011_pct_81, 
                nlcd_rowcrop_82 = iws_nlcd2011_pct_82) %>% 
  dplyr::mutate(nlcd_ag_99 = nlcd_pasture_81 + nlcd_rowcrop_82)

res <- dplyr::left_join(cdl, lg, by = c("lagoslakeid")) %>%
  tidyr::gather(key = "variable", value = "value", -lagoslakeid) %>%
  separate(variable, c("source", "variable", "code")) %>%
  group_by(lagoslakeid) %>%
  tidyr::spread(source, value) %>%
  dplyr::filter(variable != "rowcrop")

ggplot(data = res, aes(x = nlcd, y = cdl)) + 
  geom_point() +
  geom_abline(intercept = 0, slope = 1) +
  facet_wrap(~variable) +
  theme_minimal()
```