### 2018-11-02 figures/05_cafos.Rmd
infrastucture for plotting all census vars

---
output: pdf_document
header-includes: 
  - \pagenumbering{gobble}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.path = "../figures/", fig.width = 5, fig.height = 5)
source("../scripts/utils.R")
```

```{r 05_cafos, fig.height=3, echo=FALSE, eval=TRUE}
library(LAGOSNE)
library(ggplot2)
library(dplyr)
library(tidyr)
library(concaveman)
library(sf)
# setwd("data")

states <- st_as_sf(maps::map("state", fill = TRUE, plot = FALSE))
ep      <- coordinatize(readRDS("../data/ep.rds"))
ep_bbox <- concaveman(ep)$polygons
states  <- states[unlist(lapply(st_intersects(states, ep_bbox), 
                               function(x) length(x) > 0)),]

animals <- readRDS("../data/census/census.rds")
st_geometry(animals) <- NULL
animals <- left_join(ep, animals, by = "lagoslakeid")
animals <- dplyr::filter(animals, lagoslakeid != 401)

# mapview::mapview(states) +
#   mapview::mapview(animals, zcol = "hog_sales")

cowplot::plot_grid(
ggplot() +
  geom_sf(data = states) +
  geom_sf(data = animals, aes(color = hog_sales)) +
  theme(title = element_text(size = 10)) +
  labs(color = paste0("hog \n operations")),
ggplot() +
  geom_sf(data = states) +
  geom_sf(data = animals, aes(color = cattle_sales)) +
  theme(title = element_text(size = 10)) #+
  #labs(color = paste0("hog \n operations"))
)

# ---- debug ----
# library(mapview)
# llid <- 457
# t_iws <- dplyr::filter(iws, lagoslakeid == llid)
# t_county <- st_transform(county_sf(), st_crs(t_iws))
# t_county <- t_county[unlist(lapply(
#   st_intersects(t_county, t_iws), function(x) length(x) > 0)), ]
# t_raw <- animals_raw[unlist(lapply(
#   st_intersects(animals_raw, t_iws), function(x) length(x) > 0)), ]
# 
# mapview(t_raw) +
#   mapview(t_iws)
```