### 2018-11-05 figures/01_county_extent.Rmd
pull ep from 2000-2010

---
output: pdf_document
header-includes: 
  - \pagenumbering{gobble}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.path = "../figures/", fig.width = 5, fig.height = 5)
```

```{r lagos_setup, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE}
# setwd("_episodes_rmd")
library(vapour)
library(LAGOSNE)
library(magrittr)
library(dplyr)
library(lme4)
library(sf)
library(units)
library(merTools)
library(ggplot2)
library(cowplot)
library(ggridges)
library(tidyr)
library(ggExtra)

ep <- readRDS("data/ep.rds")
gdb_path <- path.expand("~/.local/share/LAGOS-GIS/lagos-ne_gis.gpkg")
# st_layers(gdb_path)
layer_name <- "COUNTY"
county <- st_read(gdb_path, layer_name, quiet = TRUE)
states <- st_as_sf(maps::map("state", fill = TRUE, plot = FALSE))

pull_ag_polygons <- function(hi_ag_iws, county){
  layer_name <- "IWS"
  crs <- st_crs(county)
  # system(paste("ogrinfo -so", gdb_path, layer_name))
  query      <- paste0("SELECT * FROM ", layer_name,
                       " WHERE lagoslakeid IN ('",
                       paste0(hi_ag_iws$lagoslakeid,
                              collapse = "', '"), "');")
  iws <- st_as_sfc(vapour_read_geometry(gdb_path, sql = query))
  iws <- st_sf(vapour_read_attributes(gdb_path, sql = query),
               geometry = iws, crs = crs)
  iws <- st_zm(iws)
  iws <- st_cast(iws, "MULTIPOLYGON") # solves error: Unknown WKB type 12

  ## filter counties that intersect hi ag iws
  hi_ag_counties_intersects <- st_intersects(county, iws)
  hi_ag_counties_intersects <- unlist(lapply(hi_ag_counties_intersects,
                                             function(x) length(x > 0)))
  county <- county[hi_ag_counties_intersects > 0,]

  ## get number of counties intersecting hi ag iws
  hi_ag_iws_intersects <- st_intersects(iws, county)
  hi_ag_iws_intersects <- unlist(lapply(hi_ag_iws_intersects,
                                        function(x) length(x > 0)))

  list(iws = iws,
       county = county,
       counties_intersects = hi_ag_counties_intersects,
       iws_intersects = hi_ag_iws_intersects)
}

get_iws_polygon <- function(lagoslakeid, crs){
  layer_name <- "IWS"
  # system(paste("ogrinfo -so", gdb_path, layer_name))
  query      <- paste0("SELECT * FROM ", layer_name,
                       " WHERE lagoslakeid IN ('",
                       paste0(lagoslakeid,
                              collapse = "', '"), "');")
  iws <- st_as_sfc(vapour_read_geometry(gdb_path, sql = query))
  iws <- st_sf(vapour_read_attributes(gdb_path, sql = query),
               geometry = iws, crs = crs)
  iws <- st_zm(iws)
  st_cast(iws, "MULTIPOLYGON") # solves error: Unknown WKB type 12
}

lg          <- lagosne_load("1.087.1")
iws_lulc    <- readRDS("data/iws_lulc.rds")
county_lulc <- readRDS("data/county_lulc.rds")

iws_vs_county_ag <- iws_lulc %>%
  left_join(dplyr::select(lg$locus, lagoslakeid, county_zoneid))  %>%
  left_join(dplyr::select(county_lulc, county_zoneid:county_ag_2011_pcent))

```

```{r mendota_oneida, echo=FALSE, message=TRUE, eval=FALSE}
llids <- lake_info(lg, 
                   name = c("Lake Mendota", "Oneida Lake"), 
                   state = c("Wisconsin", "New York"))$lagoslakeid

cat(paste(c("Lake Mendota", "Oneida Lake")[which(llids %in% ep$lagoslakeid)], "is included in the hi ag sites."))
```

<!-- * Find counties that intersect large numbers of hi ag iws with tp + tn data -->

```{r 01_county_extent, eval=TRUE, echo=FALSE, fig.width = 5.5, fig.height = 3}
# counties that overlap hi ag iws with ep data
# setwd("_episodes_rmd")

hi_ag_iws_w_ep <- pull_ag_polygons(ep, county)

n_iws <- unlist(lapply(
  st_intersects(hi_ag_iws_w_ep$county, hi_ag_iws_w_ep$iws),
  function(x) length(x)))

hi_ag_iws_w_ep$county$n_iws <- n_iws

states <- st_transform(states, st_crs(hi_ag_iws_w_ep$county))
state_intersects <- st_intersects(states, hi_ag_iws_w_ep$county)
state_intersects <- states[unlist(lapply(
  state_intersects,
  function(x) length(x) > 3)),]

hi_ag_iws_w_ep$county <- hi_ag_iws_w_ep$county[
  unlist(lapply(st_intersects(hi_ag_iws_w_ep$county, state_intersects), function(x) length(x) > 0)),]

# color by number of iws intersections
ggplot() +
  geom_sf(data = state_intersects) +
  geom_sf(data = hi_ag_iws_w_ep$county, aes(fill = n_iws)) +
  # ggtitle(paste0("# of overlappling IWS with > 40% ag & epi nutrient data \n
          # where > 4 measurements between 1995 and 2005")) +
  theme(title = element_text(size = 10)) +
  labs(fill = paste0("# lakes \n (", nrow(ep), " total)"))

# mapview::mapview(hi_ag_iws_w_ep$county, zcol = "n_iws")

# cat(paste0(nrow(ep), " total lakes"))
```