### 2019-06-17 figures/09_stream_buffer.Rmd
turn on cowplot theme

---
output: pdf_document
header-includes: 
  - \pagenumbering{gobble}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.path = "../figures/", fig.width = 7, fig.height = 7)
# setwd("figures")
source("../scripts/99_utils.R")
```

```{r 09_stream_buffer, echo=FALSE, fig.height=3, fig.width=4}
dt <- readRDS("../data/dt.rds") %>%
  mutate(diff = stream_cultivated_crops - ag_pct, 
         category = cut(diff, breaks = c(Inf, 3, -65, -Inf))) %>%
  drop_na(diff) %>%
  dplyr::select(lagoslakeid, ag_pct, stream_cultivated_crops, 
                diff, category)

# llid 6201 as a low stream ag
# llid 7026 as a high stream ag
llids <- c("6201", "7026")

ggplot(data = dt) +
  geom_point(aes(x = ag_pct, y = stream_cultivated_crops, color = category)) +
  geom_abline(aes(slope = 1, intercept = 3)) +
  geom_abline(aes(slope = 1, intercept = -65)) +
  ggrepel::geom_text_repel(data = dplyr::filter(dt, lagoslakeid %in% llids), 
            aes(x = ag_pct, y = stream_cultivated_crops, label = lagoslakeid)) + 
  theme_cowplot() +
  theme(legend.position = 0) +
  ylab("Stream buffer Ag (%)") + xlab("IWS Ag (%)") +
  xlim(c(40, 100)) + ylim(c(0, 100))
```

```{r satellite, echo=FALSE, message=FALSE, warning=FALSE}
# res_pnt  <- query_gis("LAGOS_NE_All_Lakes_4ha_POINTS", "lagoslakeid", 
#                       dplyr::filter(dt, diff > 0)$lagoslakeid)
# ggplot() + 
#   geom_sf(data = state_sf()) +
#   geom_sf(data = res_pnt)

library(curl)
library(glue)
library(slippymath)
library(rgdal)

res_iws  <- query_gis("IWS", "lagoslakeid", llids)

get_image <- function(outpath, iws){
  tile_grid <- bbox_to_tile_grid(st_bbox(
    st_transform(iws, 4326)), max_tiles = 15)

  mapbox_query_string <-
    paste0("https://api.mapbox.com/v4/mapbox.satellite/{zoom}/{x}/{y}.jpg90",
           "?access_token=",
           Sys.getenv("MAPBOX_API_KEY"))

  images <- apply(tile_grid$tiles, 1, function(i) {
    url     <- glue(mapbox_query_string, x = i[1], y = i[2], 
                zoom = tile_grid$zoom)
    outfile <- glue("../data/mapbox/{x}_{y}.jpg", x = i[1], y = i[2])
    get_if_not_exists(url, outfile)
    return(outfile)
  })

  raster_out <- compose_tile_grid(tile_grid, images)

  writeRaster(raster_out, outpath, options="INTERLEAVE=BAND")   
  return(raster_out)
}

outpaths     <- paste0("../data/mapbox/", llids, ".tif")
images       <- lapply(seq_along(outpaths), function(i) 
  get_if_not_exists(get_image, outpaths[i], raster::brick, iws = res_iws[i,]))

get_streams <- function(destfile){
  res <- lapply(seq_along(llids), function(x) 
            nhdR::nhd_plus_query(
              poly = res_iws[x,], dsn = "NHDFlowLine")$sp$NHDFlowLine)
  saveRDS(res, destfile)
  return(res)
}

# unlink("../data/nhd/res_streams.rds")
res_streams <- get_if_not_exists(get_streams, "../data/nhd/res_streams.rds")

prep_ggplot_raster <- function(r, iws){
  r_raw <- r
  r     <- projectRaster(r_raw, 
                         crs = as.character(st_crs(iws))[2])
  tab        <- as.data.frame(r, xy = TRUE)
  names(tab) <- c("x", "y", "red", "green", "blue")
  tab$red   <- scales::rescale(tab$red, to = c(0, 255))
  tab$blue  <- scales::rescale(tab$blue, to = c(0, 255))
  tab$green <- scales::rescale(tab$green, to = c(0, 255))
  tab          <- tab[complete.cases(tab),]
  tab$hex      <- rgb(tab$red, tab$green, tab$blue, maxColorValue = 255)
  tab
}

i <- 1
gg_low_ag <- ggplot() + 
  geom_raster(data = prep_ggplot_raster(images[[i]], res_iws[i,]), 
              aes(x, y, fill = I(hex))) +
  geom_sf(data = res_iws[i,], alpha = 0, color = "red") +
  geom_sf(data = res_streams[[i]], color = "cyan") +
  coord_sf(datum = NA) +
  theme_void() +
  ggtitle(llids[i])

i <- 2
gg_hi_ag <- ggplot() + 
  geom_raster(data = prep_ggplot_raster(images[[i]], res_iws[i,]), 
              aes(x, y, fill = I(hex))) +
  geom_sf(data = res_iws[i,], alpha = 0, color = "red") +
  geom_sf(data = res_streams[[i]], color = "cyan") +
  coord_sf(datum = NA) +
  theme_void() +
  ggtitle(llids[i])

cowplot::plot_grid(gg_hi_ag, gg_low_ag)
```