### 2019-06-17 figures/08_exploratory_dotplot.Rmd
rm p from tn plot and vice versa

---
output: pdf_document
header-includes: 
  - \pagenumbering{gobble}
---

```{r setup, include=FALSE, message=FALSE, echo=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.path = "../figures/", fig.width = 6.5, fig.height = 4.5)
# setwd("figures")
source("../scripts/99_utils.R")
```

```{r 08_exploratory_dotplot, echo=FALSE, message=TRUE, eval=TRUE, warning=FALSE}
pred_key <- read.csv("../data/predictor_key.csv", stringsAsFactors = FALSE)

dt       <- readRDS("../data/dt_scaled.rds")
dt_sub   <-  dt %>% 
  dplyr::select(-lagoslakeid) %>%
  # dplyr::select(tp, tn, lake_area_ha:wheat, 
  #               matches("cultivated_crops")) %>%
  # dplyr::select(-row_crop_pct, -ag_pct, -pasture_pct) %>% # rm nlcd
  # dplyr::select(-mixed.crop, -matches("other"), -forage, 
  #               -hu12_ppt_std, -water, -developed, -forest) %>% # rm junk
  # dplyr::filter(complete.cases(.)) %>%
  mutate_all(as.numeric) %>%
  janitor::remove_empty("cols") %>%
  data.frame(stringsAsFactors = FALSE)

# cor.test(dt_sub$nitrogen_atmospheric_deposition, dt_sub$tp)

dt_tp <- dt_sub %>%
  corrr::correlate() %>%
  # corrr::shave() %>%
  janitor::remove_empty(c("rows", "cols")) %>%
  mutate_if(is.numeric, round, 2) %>%
  arrange(desc(abs(tp))) %>%
  data.frame() %>%
  dplyr::select(rowname, tp, tn)

dt_tn <- dt_tp %>% 
  arrange(desc(abs(tn))) %>% 
  data.frame()

dt_tp_rows <- dt_tp$rowname
dt_tp <- dt_tp %>%
  dplyr::select(-rowname) %>%
  mutate_all(abs)
row.names(dt_tp) <- dt_tp_rows

dt_tn_rows <- dt_tn$rowname
dt_tn <- dt_tn %>%
  dplyr::select(-rowname) %>%
  mutate_all(abs)
row.names(dt_tn) <- dt_tn_rows

# ---- heatmap ----
dt_tp_hmap <- dt_sub %>%
  corrr::correlate() %>%
  # corrr::shave() %>%
  # janitor::remove_empty(c("rows", "cols")) %>%
  mutate_if(is.numeric, round, 2) %>%
  data.frame()

dt_tp_rows       <- dt_tp_hmap$rowname
dt_tp_hmap            <- dt_tp_hmap %>%
  dplyr::select(-rowname) %>%
  mutate_all(abs)
row.names(dt_tp_hmap) <- dt_tp_rows

# prettyfi variable labels
pretty_names <- pred_key$pretty[match(names(dt_tp_hmap), pred_key$var)] 
pretty_names <- pretty_names[!is.na(pretty_names)]
names(dt_tp_hmap)[which(!is.na(match(names(dt_tp_hmap), pred_key$var)))] <- 
  pretty_names
pretty_names <- pred_key$pretty[match(row.names(dt_tp_hmap), pred_key$var)] 
pretty_names <- pretty_names[!is.na(pretty_names)]
row.names(dt_tp_hmap)[which(!is.na(match(row.names(dt_tp_hmap), pred_key$var)))] <- 
  pretty_names

pheatmap::pheatmap(
  t(dt_tp_hmap),
  color = colorRampPalette(RColorBrewer::brewer.pal(n = 7, name =
                                            "Reds"))(100),
  cluster_cols = TRUE, cluster_rows = TRUE,
  na.col = "grey",
  cellheight = 7)
```

\pagebreak
```{r dotplot, message=FALSE, echo=FALSE, warning=FALSE}
# ---- dotplot ----

prep_dotplot <- function(x, col_name){
  # x <- dt_tp
  # col_name <- "tp"
  pred_key <- read.csv("../data/predictor_key.csv", stringsAsFactors = FALSE)
  res      <- x %>%
    dplyr::select(UQ(rlang::sym(col_name))) %>%
    # dplyr::filter(!is.na(UQ(rlang::sym(col_name)))) %>%
    dplyr::mutate(var = row.names(x)) %>%
    dplyr::left_join(pred_key, by = "var") %>%
    dplyr::mutate(var = factor(row.names(x), 
                            levels = rev(row.names(x))), 
                  pretty = factor(pretty, levels = rev(pretty)),
                  category = factor(category, 
                                    levels = c("Lake", "Nutrient sources",
                                               "Nutrient proxies",
                                               "Transport capacity", 
                                               "Buffer composition"))) %>%
    dplyr::filter(!(var %in% c("tp", "tn"))) %>%
    dplyr::filter(!is.na(category))
  res
}

plot_opts <- theme_cowplot() +
  theme(axis.title.y = element_blank(), 
                   axis.text = element_text(size = 8), 
                   legend.position = 0)

dt_tp <- prep_dotplot(dt_tp, "tp") %>%
  dplyr::filter(!str_detect(pretty, "^N ")) %>%
  dplyr::filter(!str_detect(pretty, " N "))

dt_tn <- prep_dotplot(dt_tn, "tn") %>%
  dplyr::filter(!str_detect(pretty, "^P ")) %>%
  dplyr::filter(!str_detect(pretty, " P "))

# head(prep_dotplot(dt_tp, "tp"))
# show_col(brewer_pal("div", "Set1")(4))                   
gg_tp <-  ggplot(data = dt_tp) + 
    geom_point(aes(x = tp, y = pretty, color = category)) + 
    scale_color_brewer(palette = "Set1") +
    xlim(0, 0.6) +
    plot_opts
gg_tn <- ggplot(data = dt_tn) + 
    geom_point(aes(x = tn, y = pretty, color = category)) +
    scale_color_brewer(palette = "Set1") +
    xlim(0, 0.6) +
    plot_opts
legend_b <- get_legend(gg_tp + 
                         theme(legend.position="bottom", 
                               legend.justification = "center", 
                               legend.text = element_text(size = 10), 
                               legend.title = element_blank()))
plot_grid(
  plot_grid(gg_tp, gg_tn), 
  legend_b, ncol = 1, rel_heights = c(1, 0.1))

```

\pagebreak
```{r densityplots, message=FALSE, echo=FALSE, warning=FALSE}
# ---- densityplots ----
library(ggforce)

ggplot(data = tidyr::gather(dt_sub)) +
  geom_density(aes(value)) +
  facet_wrap_paginate(~key, scales = "free", ncol = 2, nrow = 4, page = 1)
ggplot(data = tidyr::gather(dt_sub)) +
  geom_density(aes(value)) +
  facet_wrap_paginate(~key, scales = "free", ncol = 2, nrow = 4, page = 2)
ggplot(data = tidyr::gather(dt_sub)) +
  geom_density(aes(value)) +
  facet_wrap_paginate(~key, scales = "free", ncol = 2, nrow = 4, page = 3)

```