### 2019-07-03 figures/06_lulc_buffer_demo.Rmd
add total ag panel, closes #15

---
output: pdf_document
header-includes: 
  - \pagenumbering{gobble}
---

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.path = "../figures/", fig.width = 7, fig.height = 7)
# setwd("figures")
source("../scripts/99_utils.R")
```

```{r 06_lulc_buffer_demo, echo=FALSE, eval=TRUE, message=FALSE}
# setwd("figures")

dt <- readRDS("../data/dt.rds")

llid <- 2745
ll_pnt <- st_coordinates(
  st_transform(
    query_gis("LAGOS_NE_All_Lakes_4ha_POINTS", "lagoslakeid", llid), 4326))
ll_lake <- st_transform(
  query_gis("LAGOS_NE_All_Lakes_4ha", "lagoslakeid", llid), 4326)
network <- extract_network(lon = ll_pnt[1], 
                                   lat = ll_pnt[2], 
                                   maxsteps = Inf, 
                           buffer_dist = units::as_units(5, "km"),
                           quiet = TRUE)
network <- network[unlist(lapply(
  st_intersects(network, st_transform(ll_lake, st_crs(network))), 
  function(x) length(x) == 0)),]
iws  <- query_gis("IWS", "lagoslakeid", llid)

network_buffer <- st_buffer(network, 100)
nlcd           <- get_nlcd(template = as_Spatial(st_as_sfc(st_bbox(iws))), 
                 label = llid)
# mapview(nlcd) + mapview(iws)
network_buffer <- st_transform(network_buffer, projection(nlcd))

nlcd_df <- as.data.frame(nlcd, xy = TRUE) %>%
  dplyr::select(x, y, contains("Value"))
nlcd_df[,3] <- factor(nlcd_df[,3])
names(nlcd_df)[3] <- "value"
cols <- dplyr::filter(pal_nlcd(), code %in% unique(nlcd_df[,3]))

g_map <- ggplot() + 
  geom_raster(data = nlcd_df, aes(x = x, y = y, fill = value)) +
  scale_fill_manual(values = cols$color) +
  geom_sf(data = network_buffer) +
  coord_sf(datum = NULL) +
  theme(legend.position = "none", 
        axis.text = element_blank(), 
        axis.title = element_blank(), 
        line = element_blank(), 
        rect = element_blank(), 
        panel.grid = element_blank()) +
    ggtitle(paste0("llid: ", llid))

join_key <- function(vals){
  res <- vals %>%
    data.frame(code = ., stringsAsFactors = FALSE) %>% 
    count(code) %>%
    left_join(FedData::pal_nlcd(), by = "code") %>%
    dplyr::filter(code != 11)
  res$description <- factor(res$description, 
                            levels = res$description)
  res
}

buffer_stats <- read.csv("../data/buffer_lulc.csv") %>%
  dplyr::filter(llid == 2745) %>%
  rename(total_n_lake = total_n, 
         n_lake = n) %>%
  mutate(total_n_stream = stream_buffer_area / 1000, 
         n_stream = round((percent_stream / 100) * total_n_stream, 0))

# create a vector of nlcd codes of length n_lake or n_stream
expand_nlcd_codes <- function(x, n_col){
  x_filtered <- x %>%
    dplyr::filter(!is.na(UQ(rlang::sym(n_col))))
  Map(rep, x_filtered$code, x_filtered[,n_col])  
}

lake_buffer_stats   <- join_key(as.character(unlist(
  expand_nlcd_codes(buffer_stats, "n_lake"))))
stream_buffer_stats <- join_key(as.character(unlist(
  expand_nlcd_codes(buffer_stats, "n_stream"))))
nlcd_stats          <- join_key(as.character(values(nlcd)))

theme_opts <- theme(axis.text.x = element_blank(), 
                    legend.title = element_blank(), 
                    axis.title.x = element_blank())

g_all <- ggplot() + 
  geom_col(data = nlcd_stats, aes(x = description, 
                             y = n, 
                             fill = description)) +
  scale_fill_manual(values = nlcd_stats$color) +
  theme(legend.direction = "horizontal") +
  theme_opts +
  guides(fill = guide_legend(ncol = 2)) +
  ggtitle("Viewport LULC")

g_stream_buffer <- ggplot() + 
  geom_col(data = stream_buffer_stats, aes(x = description, 
                                  y = n, 
                                  fill = description)) +
  scale_fill_manual(values = stream_buffer_stats$color) +
  theme(legend.position = "none") +
  theme_opts +
  ggtitle("Stream buffer LULC")

g_lake_buffer <- ggplot() + 
  geom_col(data = lake_buffer_stats, aes(x = description, 
                                  y = n, 
                                  fill = description)) +
  scale_fill_manual(values = lake_buffer_stats$color) +
  theme(legend.position = "none") +
  theme_opts +
  ggtitle("Lake buffer LULC")

legend <- get_legend(g_all)

plot_grid(g_map, 
  plot_grid(g_all + theme(legend.position = "none"),
            g_lake_buffer, ggplot(),
            g_stream_buffer), 
  legend, ncol = 1)

```

```{r cdl_viewport, fig.height=5, echo=FALSE, message=FALSE}
# pull cdl lulc
cdl_key <- read.csv("../data/cdl/cdl_key.csv", stringsAsFactors = FALSE)
r  <- raster::raster("../data/cdl/2745.tif")

cdl_table <- as.data.frame(table(values(r)), stringsAsFactors = FALSE) %>%
    mutate(code = as.integer(Var1)) %>%
    dplyr::select(-Var1) %>%
    left_join(cdl_key, by = "code")

test2 <- cdl_table %>% 
  dplyr::filter(description != "Background") %>%
  rename(n = Freq) %>%
  arrange(desc(n)) %>%
  group_by(category) %>%
  summarize(group_n = sum(n, na.rm = TRUE), is_ag = first(is_ag)) %>%
  ungroup() %>%
  mutate(percent = round((group_n / sum(group_n, na.rm = TRUE)) * 100, 0)) %>%
  dplyr::filter(percent >= 1) %>%
  arrange(is_ag, percent)

# test2 <- test2 %>%
#   add_row(category = "crops", group_n = NA, is_ag = NA, percent = 
#             sum(dplyr::filter(test2, is_ag == "ag" & 
#                                 category != "pasture")$percent))

color_key <- data.frame(category = test2$category, 
                         color_cat = c("ag", "ag", "ag", "ag", "ag", "pasture", 
                                       "water", "dev", "forest"),
                         color = c(
                           RColorBrewer::brewer.pal(5, "YlOrBr"), "#DBD83D",
                           "#9ecae1", "#000000", "#2ca25f"),
                        stringsAsFactors = FALSE)
color_key <- arrange(color_key, color_cat)

test2 <- left_join(test2, color_key)
test2$category <- factor(test2$category, levels = color_key$category)

g_bar <- 
    ggplot() + 
  geom_col(data = test2, aes(x = category, 
                             y = percent, fill = category)) +
    scale_fill_manual(values = setNames(color_key$color, color_key$category)) +
  theme(legend.direction = "horizontal", 
        legend.position = 0,
        axis.text.x = element_text(angle = 90)) +
  guides(fill = guide_legend(ncol = 2)) +
  ggtitle("CDL llid = 2745")

####

cdl_df <- as.data.frame(r, xy = TRUE) %>%
  dplyr::select(x, y, contains("2745"))
names(cdl_df)[3] <- "value"
cdl_df <- left_join(cdl_df, cdl_key, by = c("value" = "code")) 
cdl_df <- dplyr::filter(cdl_df, !is.na(category), 
                        category %in% test2$category)
cdl_df$category <- factor(cdl_df$category, levels = color_key$category)

g_raster <- ggplot() + 
  geom_raster(data = cdl_df, aes(x = x, y = y, fill = category)) +
  scale_fill_manual(values = setNames(color_key$color, 
                                      color_key$category)) +
  theme_void()

####

cowplot::plot_grid(g_raster, g_bar)

```

```{r cdl_vs_nlcd, echo=FALSE, message=FALSE}
# mask nlcd lulc

test <- raster::mask(nlcd, iws) %>% 
  as.data.frame(xy = TRUE) %>%
  dplyr::select(x, y, contains("Value"))
test <- test[test[,3] %in% c(81, 82),]
test[,3] <- factor(test[,3])
names(test)[3] <- "value"
test <- test %>%
  drop_na("value")
cols     <- dplyr::filter(pal_nlcd(), code %in% unique(test[,3])) %>%
  mutate(description = c("Pasture", "Row Crops"))
cols_agg <- mutate(cols, color = "#AA7028", description = "Total Ag")[2,]

# codes 81 = pasture; 82 = row_crop

g_nlcd_agg <- ggplot() + 
  geom_raster(data = mutate(test, value = 82), aes(x = x, y = y, fill = as.factor(value))) +
  scale_fill_manual(values = cols_agg$color, labels = cols_agg$description) +
  geom_sf(data = iws, alpha = 0) +
  geom_sf(data = network) +
  geom_sf(data = ll_lake) +
  coord_sf(datum = NA) + 
  labs(fill = "") +
  theme_void() # + 
  # theme(legend.position = 0)

g_nlcd <- ggplot() + 
  geom_raster(data = test, aes(x = x, y = y, fill = value)) +
  scale_fill_manual(values = cols$color, labels = cols$description) +
  geom_sf(data = iws, alpha = 0) +
  geom_sf(data = network) +
  geom_sf(data = ll_lake) +
  coord_sf(datum = NA) + 
  labs(fill = "") +
  theme_void() # + 
  # theme(legend.position = 0)

cdl_df <- mutate(cdl_df, category = forcats::fct_collapse(cdl_df$category, 
                                                        `Corn` = "corn", 
                                                        `Soybeans` = "soybeans", 
                                                        `Pasture` = "pasture"))
g_cdl <- ggplot() + 
  geom_raster(data = dplyr::filter(cdl_df, as.character(category) %in% 
                                     c("Corn", "Soybeans", "Pasture")), 
              aes(x = x, y = y, fill = category)) +
  scale_fill_manual(values = 
                      setNames(
                        color_key$color, 
                        snakecase::to_any_case(color_key$category, "sentence"))
    ) +
  geom_sf(data = iws, alpha = 0) +
  geom_sf(data = network) +
  geom_sf(data = ll_lake) +
  coord_sf(datum = NA) + 
  labs(fill = "") +
  theme_void()

cowplot::plot_grid(
  g_nlcd_agg + ggtitle("Total Ag"), 
  g_nlcd + ggtitle("Row Crops"), 
  g_cdl + ggtitle("Individual Crops"), 
                   rel_widths = c(0.95, 1, 0.95), nrow = 1)
```

```{r raster_resolution, echo=FALSE, message=FALSE}
library(nadp)

llid  <- 2745

iws   <- query_gis("IWS", "lagoslakeid", llid)
n_dep <- nadp_load("2010", "totalN")
n_dep <- raster::crop(n_dep, as_Spatial(st_as_sfc(st_bbox(iws))))
n_dep <- mask(n_dep, iws)

# create a baseflow grid of 1000m and a ndep grid of 2338 m
gg_dep_grid <- ggplot() + 
  geom_sf(data = iws) +
  geom_sf(data = st_make_grid(iws, cellsize = 2338), alpha = 0) +
  coord_sf(datum = NA) +
  theme_void()

gg_bf_grid <- ggplot() + 
  geom_sf(data = iws) +
  geom_sf(data = st_make_grid(
    st_buffer(st_as_sfc(st_bbox(st_make_grid(iws, cellsize = 2338))), 400), 
                    cellsize = 1000), 
    alpha = 0) +
  coord_sf(datum = NA) +
  theme_void()

cowplot::plot_grid(gg_dep_grid, gg_bf_grid)
```