### 2019-07-12 figures/09_stream_buffer.Rmd
add lake and buffer outlines

---
output: pdf_document
header-includes: 
  - \pagenumbering{gobble}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.path = "../figures/", fig.width = 7, fig.height = 7)
# setwd("figures")
source("../scripts/99_utils.R")
```

```{r 09_stream_buffer, echo=FALSE, fig.height=3, fig.width=4, warning=FALSE}
dt <- readRDS("../data/dt.rds") %>%
  dplyr::filter(stream_ag > 0) %>%
  mutate(diff = stream_ag - ag, 
         category = cut(diff, breaks = c(Inf, 25, -25, -Inf))) %>%
  drop_na(diff) %>%
  dplyr::select(lagoslakeid, ag, stream_ag, 
                diff, category, iws_ha) %>%
  arrange(diff)

# first llid is a low stream ag lake
# second llid is a high stream ag lake

llids         <- c("6992", "7026")
llid_metadata <- lake_info(llids)
llid_callouts <- paste(llid_metadata$gnis_name, "-", llid_metadata$state)

ggplot(data = dt) +
  geom_point(aes(x = ag, y = stream_ag, color = category)) +
  geom_abline(aes(slope = 1, intercept = 25)) +
  # geom_abline(aes(slope = 1, intercept = 0)) +
  geom_abline(aes(slope = 1, intercept = -25)) +
  ggrepel::geom_text_repel(data = dplyr::filter(dt, lagoslakeid %in% llids), 
            aes(x = ag, y = stream_ag, label = lagoslakeid), nudge_y = 8) + 
  theme_cowplot() +
  theme(legend.position = 0) +
  ylab("Stream buffer Ag (%)") + xlab("IWS Ag (%)") +
  ylim(c(1, 100)) # + xlim(c(40, 100))

# dplyr::filter(dt, lagoslakeid %in% llids)
```

```{r satellite, echo=FALSE, message=FALSE, warning=FALSE}
# res_pnt  <- query_gis("LAGOS_NE_All_Lakes_4ha_POINTS", "lagoslakeid", 
#                       dplyr::filter(dt, diff > 0)$lagoslakeid)
# ggplot() + 
#   geom_sf(data = state_sf()) +
#   geom_sf(data = res_pnt)

library(curl)
suppressMessages(library(glue))
library(slippymath)
suppressMessages(library(rgdal))

res_iws    <- query_gis("IWS", "lagoslakeid", llids)
res_lakes  <- query_gis("LAGOS_NE_All_Lakes_4ha", "lagoslakeid", llids)

get_image <- function(outpath, iws){
  tile_grid <- bbox_to_tile_grid(st_bbox(
    st_transform(iws, 4326)), max_tiles = 15)

  mapbox_query_string <-
    paste0("https://api.mapbox.com/v4/mapbox.satellite/{zoom}/{x}/{y}.jpg90",
           "?access_token=",
           Sys.getenv("MAPBOX_API_KEY"))

  images <- apply(tile_grid$tiles, 1, function(i) {
    url     <- glue(mapbox_query_string, x = i[1], y = i[2], 
                zoom = tile_grid$zoom)
    outfile <- glue("../data/mapbox/{x}_{y}.jpg", x = i[1], y = i[2])
    get_if_not_exists(url, outfile)
    return(outfile)
  })

  raster_out <- compose_tile_grid(tile_grid, images)

  writeRaster(raster_out, outpath, options="INTERLEAVE=BAND")   
  return(raster_out)
}

outpaths     <- paste0("../data/mapbox/", llids, ".tif")
images       <- lapply(seq_along(outpaths), function(i) 
  get_if_not_exists(get_image, outpaths[i], raster::brick, iws = res_iws[i,]))

get_streams <- function(destfile){
  res <- lapply(seq_along(llids), function(x) 
            nhdR::nhd_plus_query(
              poly = res_iws[x,], dsn = "NHDFlowLine")$sp$NHDFlowLine)
  saveRDS(res, destfile)
  return(res)
}

# unlink("../data/nhd/res_streams.rds")
res_streams <- get_if_not_exists(get_streams, "../data/nhd/res_streams.rds")

prep_ggplot_raster <- function(r, iws){
  r_raw <- r
  r     <- projectRaster(r_raw, 
                         crs = as.character(st_crs(iws))[2])
  tab        <- as.data.frame(r, xy = TRUE)
  names(tab) <- c("x", "y", "red", "green", "blue")
  tab$red   <- scales::rescale(tab$red, to = c(0, 255))
  tab$blue  <- scales::rescale(tab$blue, to = c(0, 255))
  tab$green <- scales::rescale(tab$green, to = c(0, 255))
  tab          <- tab[complete.cases(tab),]
  tab$hex      <- rgb(tab$red, tab$green, tab$blue, maxColorValue = 255)
  tab
}

### draw forest outlines
# i <- 2
# forest_outline <- mapedit::editMap(
#   mapview(images[[i]])
#   )
# forest_outline <- smoothr::smooth(forest_outline$drawn)
# forest_outline <- st_intersection(
#   st_transform(forest_outline, st_crs(res_iws[i,])),
#   res_iws[i,])
# mapview(forest_outline)
# saveRDS(forest_outline, "../data/res_forest_outline_2.rds")
forest_outline_1 <- readRDS("../data/res_forest_outline_1.rds")
forest_outline_2 <- readRDS("../data/res_forest_outline_2.rds")

i <- 1
gg_low_ag <- ggplot() + 
  geom_raster(data = prep_ggplot_raster(images[[i]], res_iws[i,]), 
              aes(x, y, fill = I(hex))) +
  geom_sf(data = res_iws[i,], alpha = 0, color = "red") +
  geom_sf(data = res_lakes[i,], alpha = 0, color = "cyan") +
  geom_sf(data = forest_outline_1, color = "green3") +
  geom_sf(data = st_difference(
      st_transform(res_streams[[i]], st_crs(res_lakes[i,])), 
      res_lakes[i,]), 
    color = "cyan") +
  coord_sf(datum = NA) +
  theme_void() +
  ggtitle(llid_callouts[i])

i <- 2
gg_hi_ag <- ggplot() + 
  geom_raster(data = prep_ggplot_raster(images[[i]], res_iws[i,]), 
              aes(x, y, fill = I(hex))) +
  geom_sf(data = res_iws[i,], alpha = 0, color = "red") +
  geom_sf(data = res_lakes[i,], alpha = 0, color = "cyan") +
  geom_sf(data = forest_outline_2, color = "green3") +
  geom_sf(data = st_difference(
      st_transform(res_streams[[i]], st_crs(res_lakes[i,])), 
      res_lakes[i,]), 
    color = "cyan") +
  coord_sf(datum = NA) +
  theme_void() +
  ggtitle(llid_callouts[i])

cowplot::plot_grid(gg_hi_ag, gg_low_ag)
```