### 2019-10-22 tables/01_predictors.Rmd
pull iqr

---
output:
  pdf_document
header-includes: 
  - \pagenumbering{gobble}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, fig.path = "../figures/", fig.width = 5, fig.height = 6)
# setwd("figures")
```

```{r summary_table, eval=TRUE}
library(kableExtra)
suppressMessages(library(dplyr))
library(magrittr)

pred_key <- read.csv("../data/predictor_key.csv", 
                     stringsAsFactors = FALSE)

cdl_vars <- readRDS("../data/dt.rds") %>%
  dplyr::select(ag:wheat) %>%
  summarize_all(median, na.rm = TRUE) %>%
  tidyr::gather() %>%
  dplyr::filter(value >= 1.4) %>%
  arrange(desc(value)) %>%
  pull(key)

# setwd("tables")
dt <- readRDS("../data/dt.rds") %>%
  dplyr::select(tp:clay_pct, -matches("zoneid"), -matches("pct"), -matches("input"), cdl_vars) %>%
  tidyr::gather(key = "variable") %>%
  group_by(variable) %>%
  mutate(value = as.numeric(value)) %>%
  mutate(q10 = quantile(value, probs = c(0.1), na.rm = TRUE),
         q25 = quantile(value, probs = c(0.25), na.rm = TRUE),
         q50 = quantile(value, probs = c(0.5), na.rm = TRUE),
         q75 = quantile(value, probs = c(0.75), na.rm = TRUE),
         q90 = quantile(value, probs = c(0.9), na.rm = TRUE), 
         ratio = q90 / q10, 
         n = sum(!is.na(value), na.rm = TRUE)) %>%
  dplyr::distinct(signif(q10, 2), signif(q25, 2), signif(q50, 2), signif(q75, 2), signif(q90, 2), ratio, n) %>%
  dplyr::select(variable, 
         q10 = `signif(q10, 2)`,
         q25 = `signif(q25, 2)`,
         q50 = `signif(q50, 2)`,
         q75 = `signif(q75, 2)`,
         q90 = `signif(q90, 2)`,
         q10_q90_ratio = ratio,
         n) %>%
  left_join(readRDS("../data/dt_units.rds"), by = "variable") %>%
  dplyr::select(variable, units, everything())

dt <- ungroup(dt)
dt <- left_join(dt, pred_key, by = c("variable" = "var"))
dt <- dplyr::filter(dt, !is.na(pretty))

dt <- mutate(dt, pretty = case_when(
  nchar(units) > 0 ~ paste0(pretty, " (", units, ")"),
  TRUE ~ pretty), 
  IQR = paste0(q25, " - ", q75))
dt <- dplyr::select(dt, Variable = pretty, 
                    Median = q50, Q25 = q25, Q75 = q75)

knitr::kable(dt, digits = 0, format = "markdown") 
#%>%
 # add_header_above(c(" ", " ", Ranges = 3, " ", " "))
```