### 2019-10-23 tables/02_cdl_key.Rmd

---
output:
  pdf_document
header-includes: 
  - \pagenumbering{gobble}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, fig.path = "../figures/", fig.width = 5, fig.height = 6)
```

```{r cdl_key, eval=TRUE}
library(kableExtra)
library(dplyr)
library(magrittr)

# setwd("tables")
# ---- cdl_key ----
cdl_key <- read.csv("../data/cdl/cdl_key.csv", stringsAsFactors = FALSE)
cdl_key %>%
  group_by(category) %>%
  sample_n(1) %>%
  arrange(is_ag, code) %>%
  dplyr::select(-is_forage) %>%
  dplyr::select(category, ag = is_ag, example = description, code) %>%
  knitr::kable(digits = 0, format = "markdown", caption = "Example of CDL categorizations.") 
```

\pagebreak
```{r cdl_summary}
# setwd("tables")
# ---- cdl summary ----
cdl_data <- read.csv("../data/cdl/cdl_summary.csv", stringsAsFactors = FALSE)

dt <- apply(cdl_data, 2, function(x) quantile(x, c(0.05, 0.5, 0.95), 
                                        na.rm = TRUE)) %>%
  data.frame() %>%
  dplyr::select(-llid)

dt %>%
  t() %>%
  knitr::kable(caption = "CDL summary statistics", digits = 2,
               format = "latex", booktabs = TRUE) %>%
  kable_styling(latex_options = c("hold_position"), full_width = FALSE) %>%
  add_header_above(c(" ", "Quantiles" = 3))
```