### 2019-11-27 figures/11_map.Rmd
red cached simple hu4s

---
output: pdf_document
header-includes: 
  - \pagenumbering{gobble}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.path = "../figures/", fig.width = 6.5, fig.height = 4.5)
# setwd("figures")
source("../scripts/99_utils.R")
```

```{r 11_map, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE, results='hide'}
# cl            <- cart_layers()
# ep            <- cl$ep
# states        <- cl$states
# states_shaded <- cl$states_shaded
# study_bbox    <- cl$study_bbox
study_bbox    <- st_read("../data/gis.gpkg", "study_bbox", quiet = TRUE)[3,]
ep            <- st_read("../data/gis.gpkg", "ep", quiet = TRUE)
states_all    <- st_read("../data/gis.gpkg", "states_all", quiet = TRUE)
states_focal  <- st_read("../data/gis.gpkg", "states_focal", quiet = TRUE)
hu4_focal     <- st_read("../data/gis.gpkg", "hu4s_focal_simple", quiet = TRUE)

# outline <- ne_countries(country = c("canada", "mexico", "united states of america"), 
#                      returnclass = "sf")
# inset_bbox <- mapedit::drawFeatures()
# sf::st_write(inset_bbox, "../data/inset_bbox.gpkg", 
#              layer_options = "OVERWRITE=YES")
# inset_bbox <- st_read("../data/inset_bbox.gpkg")
# outline <- st_crop(outline, inset_bbox)
# outline <- outline %>%
#   st_transform(LAGOSNEgis:::albers_conic()) %>%
#   st_snap(x = ., y = ., tolerance = 0.000001) %>%
#   st_union()

gg_states <- 
  ggplot() + 
    geom_sf(data = study_bbox, alpha = 1, size = 0.6, 
          color = "black", fill = "#1D9DCF") +
    geom_sf(data = states_all, fill = "gray", color = "black", size = 0.3) +
    # geom_sf(data = outline, color = "black", alpha = 0, size = 0.8) +
    # geom_sf(data = hu4, fill = "grey") +
    geom_sf(data = states_focal, fill = "white", color = "black", size = 0.4) + 
    geom_sf(data = ep, size = 0.9) +
    coord_sf(datum = NA) +
    theme_void()

gg_hu4 <- 
  ggplot() + 
      geom_sf(data = study_bbox, alpha = 1, size = 0.6, 
              color = "black", fill = "#1D9DCF") +
      geom_sf(data = states_all, fill = "gray", color = "black", size = 0.3) +
      geom_sf(data = hu4_focal, fill = "white", color = "black", size = 0.4) +
      # geom_sf(data = ep, size = 1) +
      coord_sf(datum = NA) +
      theme_void() + 
      theme(legend.position = 0)

cowplot::plot_grid(gg_states, gg_hu4, ncol = 2, labels = "AUTO", vjust = 8, hjust = -1)
```