### 2019-12-02 figures/07_model-selection.Rmd
fix global effect handling

---
output: pdf_document
header-includes: 
  - \pagenumbering{gobble}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.path = "../figures/", eval = TRUE, warning = FALSE, message = FALSE)
# setwd("figures")
source("../scripts/99_utils.R")
```

```{r gis_setup}
dt      <- readRDS("../data/dt.rds")
ids     <- readRDS("../data/dt.rds") %>%
  group_by(hu4_zoneid) %>% 
  tally()

hu4 <-  LAGOSNEgis::query_gis("HU4", "ZoneID", ids$hu4_zoneid) %>%
  st_cast(to = "MULTIPOLYGON") %>%
  left_join(ids, by = c("ZoneID" = "hu4_zoneid"))
n_cat  <- 6
colors <- c(#"#FFFFFF", 
            viridisLite::viridis(n_cat))
hu4    <- mutate(hu4, 
            n_cat = cut(hu4$n, breaks = 
                round(classInt::classIntervals(hu4$n, n = n_cat,
                                               style = "kmeans")$brks, 0)
                ))

hu4 <- dplyr::filter(hu4, !is.na(n_cat))
hu4 <- dplyr::left_join(hu4, 
                        distinct(dplyr::select(dt, contains("hu4_")), 
                                 hu4_zoneid, .keep_all = TRUE), 
                        by = c("ZoneID" = "hu4_zoneid"))

re_brms        <- readRDS("../data/mcmc/re_brms.rds")
pred_key       <- read.csv("../data/predictor_key.csv", stringsAsFactors = FALSE)
```

```{r re_global}
# tp, tn
# ag, row_crop, pasture, corn, soybeans

re_global          <- dplyr::rbind_all(lapply(re_brms, function(x) x$re_global))
re_global$term     <- unlist(lapply(re_brms, 
                                function(x) get_re_text(as.character(x$formula)[1])))
re_global$response <- c(rep("tp", nrow(re_global) / 2), rep("tn", nrow(re_global) / 2))
re_vars         <- paste0("b_", re_global$term)
re_global_l     <- lapply(seq_len(length(re_brms)), function(i){
  spread_draws(re_brms[[i]], (!!sym(re_vars[i]))) %>% 
    summarize(value = list(tibble::enframe(
    quantile((!!sym(re_vars[i])), probs=c(0.05,0.5,0.95))))) %>% unnest() %>%
      tidyr::spread(name, value)
})
re_global[,1:3] <- dplyr::bind_rows(re_global_l)

re_global          <- left_join(re_global, pred_key, by = c("term" = "var")) %>%
  dplyr::select(pretty, `5%`:`95%`, term, response)
re_global <- dplyr::filter(re_global, term != "rowvcropvpct")
# group_by(re_global, response) %>%
#   arrange(desc(`50%`))
re_names <- unlist(lapply(re_brms, function(x) 
  get_re_text(as.character(x$formula)[[1]])))[1:length(re_brms)]
re_brms <- re_brms[re_names != "rowvcropvpct"]
```

```{r pre_tp_fe, out.width="3.45in", out.height="4in"}
# autoselect best model
re_names <- unlist(lapply(re_brms, function(x) 
  get_re_text(as.character(x$formula)[[1]])))[1:(length(re_brms)/2)]
best_tp <- loo_compare(
  lapply(seq_len(length(re_brms)/2), function(x) re_brms[[x]]$loo))
best_tp <- which(re_names == row.names(best_tp)[1])

fit_tp         <- re_brms[[best_tp]] 
signif_hu4_ids <- fit_tp$re_signif_ids 
signif_hu4     <- dplyr::filter(hu4, ZoneID %in% signif_hu4_ids)

# plot tp model fixed effects summary
# summary(fit_tp)

b_vars <- stringr::str_subset(
  tidybayes::get_variables(fit_tp), 
  c("^(b_)(?!.*Intercept).*$"))

gg_tp <- spread_draws(fit_tp, `b_.*`, regex = TRUE) %>%
  tidyr::gather(key = "key", value = "value", -.chain, -.iteration, -.draw) %>%
  dplyr::filter(key != "b_Intercept") %>%
  mutate(key = stringr::str_replace(key, "b_", "")) %>%
  left_join(pred_key, by = c("key" = "varv")) %>%
  group_by(pretty) %>%
  summarize(value = list(tibble::enframe(
    quantile(value, probs=c(0.05,0.5,0.95))))) %>% unnest() %>%
  tidyr::spread(name, value) %>%
  dplyr::bind_rows(
    dplyr::select(re_global[best_tp,], -term, -response), 
    data.frame(pretty = "", `5%` = "", `50%` = "", `95%` = "", 
               stringsAsFactors = FALSE)
    ) %>%
  rowwise() %>%
  mutate(signif = (`5%` > 0 & `95%` > 0)| (`5%` < 0 & `95%` < 0)) %>%
  left_join(dplyr::select(pred_key, pretty, category)) %>%
  mutate(category = tidyr::replace_na(category, "Land-use cover")) %>%
  mutate(category = factor(category, 
                           levels = rev(c("Nutrient inputs", "Nutrient transport", 
                                      "Buffer configuration", "Lake", 
                                      "Land-use cover")))) %>%
  mutate(category = forcats::fct_recode(category, 
                                        Buffer = "Buffer configuration", 
                                        LULC = "Land-use cover")) %>%
  arrange(category) %>%
  ungroup() %>% distinct(pretty, .keep_all = TRUE) %>%
  mutate(pretty = factor(pretty, levels = pretty)) %>%
  arrange(category, pretty)
```

```{r pre_tn_fe, out.width="3.45in", out.height="4in"}
# autoselect best tn model
re_names <- unlist(lapply(re_brms, function(x) 
  get_re_text(as.character(x$formula)[[1]])))[seq((length(re_brms)/2) + 1,
                                                  length(re_brms))]
best_tn <- loo_compare(
  lapply(seq((length(re_brms)/2) + 1, length(re_brms)), 
         function(x) re_brms[[x]]$loo))
best_tn <- which(re_names == row.names(best_tn)[1]) + (length(re_brms)/2)

fit_tn         <- re_brms[[best_tn]] 
signif_hu4_ids <- fit_tn$re_signif_ids 
signif_hu4     <- dplyr::filter(hu4, ZoneID %in% signif_hu4_ids)

# plot tn model fixed effects summary
# summary(fit_tn)
gg_tn <- spread_draws(fit_tn, `b_.*`, regex = TRUE) %>%
  tidyr::gather(key = "key", value = "value", -.chain, -.iteration, -.draw) %>%
  dplyr::filter(key != "b_Intercept") %>%
  mutate(key = stringr::str_replace(key, "b_", "")) %>%
  left_join(pred_key, by = c("key" = "varv")) %>%
  group_by(pretty) %>%
  summarize(value = list(tibble::enframe(
    quantile(value, probs=c(0.05,0.5,0.95))))) %>% unnest() %>%
  tidyr::spread(name, value) %>%
  dplyr::bind_rows(
    dplyr::select(re_global[best_tn,], -term, -response), 
    data.frame(pretty = "", `5%` = "", `50%` = "", `95%` = "", 
               stringsAsFactors = FALSE)
    ) %>%
  rowwise() %>%
  mutate(signif = (`5%` > 0 & `95%` > 0)| (`5%` < 0 & `95%` < 0)) %>%
  left_join(dplyr::select(pred_key, pretty, category)) %>%
  mutate(category = tidyr::replace_na(category, "Land-use cover")) %>%
  mutate(category = factor(category, 
                           levels = rev(c("Nutrient inputs", "Nutrient transport", 
                                      "Buffer configuration", "Lake", 
                                      "Land-use cover")))) %>%
  mutate(category = forcats::fct_recode(category, 
                                        Buffer = "Buffer configuration", 
                                        LULC = "Land-use cover")) %>%
  arrange(category) %>%
  ungroup() %>% distinct(pretty, .keep_all = TRUE) %>%
  mutate(pretty = factor(pretty, levels = pretty)) %>%
  arrange(category, pretty)
```

```{r fe_axis}
# specify shared x-axis range
x_max <- c(gg_tp$`95%`, gg_tn$`95%`)[which.max(c(gg_tp$`95%`, gg_tn$`95%`))]
x_min <- c(gg_tp$`5%`, gg_tn$`5%`)[which.min(c(gg_tp$`5%`, gg_tn$`5%`))]
```

```{r tp_fe, out.width="3.45in", out.height="4in"}
(gg_tp_fe <- fe_dotplot(gg_tp, "left", cat_labels = TRUE,
                        x_min, x_max, "TP"))
```

```{r tn_fe, out.width="3.45in", out.height="4in"}
(gg_tn_fe <- fe_dotplot(gg_tn, "right", cat_labels = FALSE, 
                        x_min, x_max, "TN"))
```

```{r fe}
cowplot::plot_grid(
cowplot::plot_grid(
  equalize_panels(
    gg_tp_fe + theme(axis.title.x = element_blank())
  ), 
  equalize_panels(
    gg_tn_fe + 
      theme(axis.text.y = element_blank(), 
            axis.title.x = element_blank())), 
  rel_widths = c(1, 0.95)),

ggplot(data.frame(x = 1, y = 1, label = "Standardized coefficients", stringsAsFactors = FALSE)) + 
  geom_text(aes(x = x, y = y, label = label)) +
  theme_void(),
ncol = 1, rel_heights = c(1, 0.05))
```

```{r tp_re, fig.width=3.25, fig.height=4.5}
fit_tp$re$hu4vzoneid <- gsub("HU4_", "", fit_tp$re$hu4vzoneid)
fit_tp$re <- fit_tp$re %>%
  mutate_at(vars(matches("%")), function(x) x +
              as.numeric(fit_tp$re_global["50%"]))

gg_tp_re <- ggplot(data = fit_tp$re) +
  geom_pointrange(aes(x = hu4vzoneid, y = `50%`, 
                      ymin = `5%`, ymax = `95%`, color = signif)) +
  scale_color_manual(values = c("FALSE" = "black", "TRUE" = "red")) +
  geom_hline(yintercept = as.numeric(fit_tp$re_global["50%"])) +
  coord_flip() +
  cowplot::theme_cowplot() +
  theme(rect = element_rect(fill = "transparent")) +
  ggtitle("TP") + xlab("Zone ID") +
  theme(plot.title = element_text(hjust = 0.5, face = "bold")) +
  theme(legend.position = 0)

gg_tp_re
```

```{r tn_re_hu4, fig.width = 3.77, fig.height = 4.5}
# setwd("figures")
study_bbox    <- st_read("../data/gis.gpkg", "study_bbox", quiet = TRUE)[3,]
states_all    <- st_read("../data/gis.gpkg", "states_all", quiet = TRUE)
hu4_all       <- st_read("../data/gis.gpkg", "hu4s_focal_simple", quiet = TRUE)

# map significant random effects hu4s
gg_hu4 <- 
  ggplot() +
    geom_sf(data = study_bbox, alpha = 1, size = 0.6, 
              color = "black", fill = "#1D9DCF") +
    geom_sf(data = states_all, fill = "gray", color = "black", size = 0.3) +
    geom_sf(data = hu4_all, fill = "white", color = "black", size = 0.4) +
    geom_sf(data = signif_hu4, color = "red", alpha = 0) +
    # geom_text(data = coords, aes(x = x, y = y, label = ZoneID)) +   
    coord_sf(datum = NA) +
    theme_minimal() + 
    theme(legend.position = "bottom", 
          axis.title = element_blank(), 
          legend.title = element_blank()) + 
    scale_fill_manual(values = colors, levels(hu4$n_cat)) + 
    guides(guide_legend(nrow = 2))  # +
  # ggtitle("Ag sensitive regions")

gg_hu4
```

```{r tn_re, fig.width=3.25, fig.height=4.5}
# dotplot tn hu4 random effects
fit_tn$re$hu4vzoneid <- gsub("HU4_", "", fit_tn$re$hu4vzoneid)
fit_tn$re <- fit_tn$re %>%
  mutate_at(vars(matches("%")), function(x) x + 
              as.numeric(fit_tn$re_global["50%"]))

gg_tn_re <- ggplot(data = fit_tn$re) +
  geom_hline(yintercept = as.numeric(fit_tn$re_global["50%"])) +
  geom_pointrange(aes(x = hu4vzoneid, y = `50%`, 
                      ymin = `5%`, ymax = `95%`, color = signif)) +
  scale_color_manual(values = c("FALSE" = "black", "TRUE" = "red")) +
  coord_flip() +
  cowplot::theme_cowplot() +
  theme(rect = element_rect(fill = "transparent")) +
  ggtitle("TN") + xlab("Zone ID") +
  theme(plot.title = element_text(hjust = 0.5, face = "bold")) +
  theme(legend.position = 0)

gg_tn_re
```

```{r tn_re_compare, fig.height=3, fig.width=4}
# compare lakes in signif hu4s
dt        <- readRDS("../data/dt_scaled.rds") %>%
  mutate(signif_hu4 = hu4_zoneid %in% signif_hu4_ids) %>%
  tidyr::gather(key = "key", value = "value", 
                      -lagoslakeid, -hu4_zoneid, -hu12_zoneid, -signif_hu4) %>%
  group_by(key, signif_hu4) %>%
  mutate(median = median(value, na.rm = TRUE))

dt_predictors <- dplyr::filter(dt, key %in% 
                                 c(#"tp",
                                   "nitrogen_atmospheric_deposition",
                                   "clay_pct")) %>%
  dplyr::left_join(pred_key, by = c("key" = "var"))
dt_predictors[is.na(dt_predictors$pretty), "pretty"] <- 
  sapply(dt_predictors[is.na(dt_predictors$pretty), "key"], toupper)

dt_modelled <- dplyr::filter(dt, key %in% 
                               c("maxdepth", "tn", "soil_org_carbon",
                                 "buffer_cultivated_crops", "corn",
                                 "nitrogen_fertilizer_use"))

gg_hist <- ggplot(data = dt_predictors) + 
  geom_histogram(aes(x = value, fill = signif_hu4), alpha = 0.7) +
  geom_vline(aes(xintercept = median, color = signif_hu4), 
             linetype = "dashed") +
  scale_fill_manual(values = c("FALSE" = "black", "TRUE" = "red")) +
  scale_color_manual(values = c("FALSE" = "black", "TRUE" = "red")) +
  cowplot::theme_cowplot() +
  theme(legend.position = 0) +
  theme(rect = element_rect(fill = "transparent"), 
        strip.background = element_rect(fill = "transparent"), 
        strip.text = element_text(face = "bold")) +
  facet_wrap(~pretty) +
  xlab("Standardized value")

# line graph tn vs tp
dt        <- readRDS("../data/dt_scaled.rds") %>%
  mutate(signif_hu4 = hu4_zoneid %in% signif_hu4_ids) %>%
  tidyr::gather(key = "key", value = "value", 
                      -lagoslakeid, -hu4_zoneid, -hu12_zoneid, -signif_hu4, -tn) %>%
  dplyr::filter(key %in% c(c("tp"))) %>%
  tidyr::spread(key, value)

gg_tp <- ggplot(data = dt) + 
  geom_point(aes(x = tp, y = tn, color = signif_hu4)) +
  geom_line(stat = "smooth", method = lm, se = FALSE, 
            aes(x = tp, y = tn, color = signif_hu4)) +
  scale_fill_manual(values = c("FALSE" = "black", "TRUE" = "red")) +
  scale_color_manual(values = c("FALSE" = "black", "TRUE" = "red")) +
  cowplot::theme_cowplot() +
  theme(legend.position = 0) + 
  theme(rect = element_rect(fill = "transparent"))

plot_grid(gg_hist, nrow = 1)
# plot_grid(gg_hist, gg_tp, nrow = 2)

lg <- lagosne_load("1.087.1")
gg_line <- fit_tn$re %>%
  ungroup() %>%
  mutate(hu4vzoneid = paste0("HU4_", hu4vzoneid)) %>%
  left_join(dplyr::select(lg$hu4.chag, 
                          hu4_zoneid, hu4_dep_totaln_2010_mean), 
            by = c("hu4vzoneid" = "hu4_zoneid")) %>%
ggplot() +
  geom_pointrange(aes(x = hu4_dep_totaln_2010_mean, y = `50%`,
                      ymin = `5%`, ymax = `95%`, color = signif)) +
  scale_color_manual(values = c("FALSE" = "black", "TRUE" = "red")) +
  theme_cowplot() +
  ylab("TN model, corn RE slope") + xlab("N deposition") +
  cowplot::theme_cowplot() +
  theme(legend.position = 0) + 
  theme(rect = element_rect(fill = "transparent"))

gg_tn_re

# gg_line

```

```{r re}
cowplot::plot_grid(
cowplot::plot_grid(gg_tp_re + theme(axis.title.x = element_blank(), 
                                     plot.margin = unit(c(0,0,0,0), "cm")),
                   gg_tn_re + theme(axis.text.y = element_blank(), 
                                    axis.title.y = element_blank(), 
                                    axis.title.x = element_blank(),
                                    plot.margin = unit(c(0,0,0,0), "cm")), 
                   rel_widths = c(1, 0.9)), 
ggplot(data.frame(x = 0, y = 0, label = "Standardized coefficients", stringsAsFactors = FALSE)) + 
  geom_text(aes(x = x, y = y, label = label), size = 5) +
  theme_void() + 
  theme(plot.margin = unit(c(0,0,0,0), "cm")),
ncol = 1, rel_heights = c(1, 0.05))

cowplot::plot_grid(gg_tp_re, 
                   gg_tn_re + 
                     theme(axis.text.y = element_blank(), 
                           axis.title.y = element_blank()) + 
                     gghighlight(max(`50%`), n = 1), 
                   rel_widths = c(1, 0.9))
```

```{r re-comparison, fig.width=4}
re_global <- dplyr::filter(re_global, term != "rowvcropvpct")
gg_tn <- ggplot(data = dplyr::filter(re_global, response == "tn")) +
    geom_pointrange(aes(x = term, y = `50%`, 
                      ymin = `5%`, ymax = `95%`), color = "red") +
  geom_hline(yintercept = 0, linetype = "dashed") +
  expand_limits(y = c(-0.2, 0.5)) +
  coord_flip() +
  theme(axis.title.y = element_blank(),
        axis.title.x = element_blank(),
        strip.placement = "outside", 
        strip.background = element_blank(), 
        panel.background = element_rect(fill = "transparent", color = NA),
        plot.background = element_rect(fill = "transparent", color = NA), 
        panel.grid = element_blank(),
        rect = element_rect(fill = "transparent"), 
        axis.line = element_line(), 
        axis.text.y = element_blank()) +
  ggtitle("TN") + 
  theme(plot.title = element_text(hjust = 0.5, face = "bold"))

gg_tp <- ggplot(data = dplyr::filter(re_global, response == "tp")) +
    geom_pointrange(aes(x = term, y = `50%`, 
                      ymin = `5%`, ymax = `95%`), color = "red") +
  geom_hline(yintercept = 0, linetype = "dashed") +
  expand_limits(y = c(-0.2, 0.5)) +
  coord_flip() +
  theme(axis.title.y = element_blank(),
        axis.title.x = element_blank(),
        strip.placement = "outside", 
        strip.background = element_blank(), 
        panel.background = element_rect(fill = "transparent", color = NA),
        plot.background = element_rect(fill = "transparent", color = NA), 
        panel.grid = element_blank(),
        rect = element_rect(fill = "transparent"), 
        axis.line = element_line()) +
  ggtitle("TP") + 
  theme(plot.title = element_text(hjust = 0.5, face = "bold"))

cowplot::plot_grid(
cowplot::plot_grid(gg_tp, gg_tn, rel_widths = c(1, 0.75)),
ggplot(data.frame(x = 1, y = 1, label = "Standardized coefficients", stringsAsFactors = FALSE)) + 
  geom_text(aes(x = x, y = y, label = label)) +
  theme_void(),
ncol = 1, rel_heights = c(1, 0.05))
```

```{r tn_interaction, eval=FALSE}
# lapply(re_brms, function(x) x$formula) 
fit <- re_brms[[6]]
fit$formula
head(fit$res_med)
test <- readRDS("../data/dt_scaled.rds") %>%
  dplyr::select(lagoslakeid, hu4_nitrogen_atmospheric_deposition, corn, 
                maxdepth, soil_org_carbon, 
                nitrogen_fertilizer_use, buffer_cultivated_crops) %>%
  left_join(fit$res_med)

cor.test(test$.residual_median, test$hu4_nitrogen_atmospheric_deposition * test$soil_org_carbon)
abline(0, 1)
```

```{r 07_model_selection, fig.height=3, echo=FALSE, fig.width=6, eval=FALSE}
# gg_hu4_ndep <- ggplot() +
#     geom_sf(data = hu4, aes(fill = hu4_nitrogen_atmospheric_deposition)) +
#     geom_sf(data = signif_hu4, color = "red", alpha = 0) +
#     theme_minimal() + 
#     theme(legend.position = "bottom", 
#           axis.title = element_blank(), 
#           legend.title = element_blank()) +
#   ggtitle("N Deposition")

# gg_hu4_clay <- ggplot() +
#     geom_sf(data = hu4, aes(fill = hu4_clay_pct)) +
#     theme_minimal() + 
#     theme(legend.position = "bottom", 
#           axis.title = element_blank(), 
#           legend.title = element_blank())

# hist(hu4$n, n = 20)
# table(as.character(dt$hu4_zoneid))

# hist(as.vector(table(dt$hu4_zoneid)))  
# good_hu4s <- group_by(dt, hu4_zoneid) %>% 
#   tally() %>%
#   dplyr::filter(n > 220)  
# dt <- dplyr::filter(dt, hu4_zoneid %in% good_hu4s$hu4_zoneid)
# dt <- dplyr::filter(dt, tp < 500 & chla < 150)

# create alpha sample size weightings
# dt <- left_join(dt, ungroup(count(group_by(dt, hu4_zoneid))), 
#                 by = "hu4_zoneid") %>%
#   dplyr::filter(n >= 3) %>%
#   mutate(n = n / max(n, na.rm = TRUE))

# join iws ag
# dt <- left_join(dt, ungroup(dplyr::select(mutate(
#   group_by(lg$iws.lulc, lagoslakeid), 
#   ag_pct = sum(iws_nlcd2011_pct_81,
#                            iws_nlcd2011_pct_82, na.rm = TRUE)), 
#   lagoslakeid, ag_pct))
#   )

# create hu4 labels
# centroid_xy <- sf::st_as_text(sf::st_geometry(sf::st_centroid(hu4[,"ZoneID"])))
# extract_coords <- function(messy_centroid){
#   res <- stringr::str_split(messy_centroid, "\\(", simplify = TRUE)[2]
#   res <- stringr::str_split(res, "\\)", simplify = TRUE)[1]
#   stringr::str_split(res, " ", simplify = TRUE)
# }
# coords <- data.frame(matrix(
#             as.numeric(unlist(lapply(centroid_xy, extract_coords))),
#             ncol = 2, byrow = TRUE), stringsAsFactors = FALSE)
# names(coords) <- c("x", "y")
# coords$ZoneID <- gsub("HU4_", "", hu4[,"ZoneID"]$ZoneID)
# 
# gg_smooth <- ggplot(data = dt, 
#                     aes(x = ag_pct, y = tp, color = hu4_zoneid, alpha = n)) + 
#     geom_line(stat = "smooth", method = lm, se = FALSE, aes(alpha = n))  +
#     geom_smooth(aes(group = 1), method = lm, se = FALSE, 
#                 color = "black", linetype = "dashed") +
#     # geom_abline(slope = 1.55, intercept = -19.86, color = "red") +
#     # scale_y_log10() + scale_x_log10() +
#     theme(legend.position = "") +
#     labs(x = "Ag. LULC (pct, NLCD 2011)", 
#          y = "TP")  +
#     scale_y_continuous(breaks = seq(0, 250, by = 50), limits = c(0, 300))
#     
# plot_grid(gg_hu4_n, gg_hu4_ndep)
``` 
```{r prediction_lines, echo=FALSE, eval=FALSE}
# plot prediction lines
# corn vs tn

# hu4_signif$ZoneID
dt        <- readRDS("../data/dt_scaled.rds") 
names(dt) <- gsub("_", "v", names(dt))

hu4_id               <- unique(dt$hu4vzoneid)
conditions           <- data.frame(hu4vzoneid = hu4_id)
rownames(conditions) <- hu4_id

me_zone <- marginal_effects(fit_tn, conditions = conditions, 
                            re_formula = NULL, method = "predict")
plot(me_zone, points = TRUE)

n_predict <- 3
reps      <- length(unique(dt$hu4vzoneid)) / 3
reps      <- 2
new_data <- data.frame(
      maxdepth = rep(modelr::seq_range(dt$maxdepth, n = n_predict), reps), 
      soilvorgvcarbon = rep(modelr::seq_range(dt$maxdepth, n = n_predict), reps), 
      nitrogenvfertilizervuse = rep(modelr::seq_range(dt$nitrogenvfertilizervuse, 
                                                  n = n_predict), reps), 
      buffervcultivatedvcrops = rep(modelr::seq_range(dt$buffervcultivatedvcrops, 
                                                  n = n_predict), reps), 
      corn = rep(modelr::seq_range(dt$corn, n = n_predict), reps), 
      hu4vzoneid = unique(dt$hu4vzoneid)[1:2],
      stringsAsFactors = FALSE)
new_data <- expand.grid(new_data)

test <- add_predicted_draws()

test <- predict(fit_tn, newdata = new_data, allow_new_levels = TRUE)
test2 <- cbind(new_data, test)
plot(test2$corn, test2$Estimate)

ran_ef <- as.data.frame(ranef(fit_tn)$hu4vzoneid) %>%
  mutate(zoneid = rownames(.)) %>%
  arrange(desc(Estimate.corn))
%>%
  dplyr::filter(zoneid == hu4_id)

# plot(me_zone, ncol = 1)$agvpct # + 
  # geom_abline(intercept = 0, slope = ran_ef$Estimate.bagvpct_Intercept + 
  #               broom::tidy(fit)$estimate[3]) + 
  # ggtitle(hu4_id)

# ggplot() + 
#     geom_point(data = dplyr::filter(dt, 
#                                   hu4vzoneid == hu4_id), 
#              aes(x = agvpct, y = tp)) + 
#   geom_abline(intercept = 0, slope = ran_ef$Estimate.bagvpct_Intercept + 
#                 broom::tidy(fit)$estimate[3]) +
#   ylim(0, 450) + xlim(40, 90) + ggtitle(hu4_id)

dplyr::filter(dt, hu4vzoneid %in% signif_hu4$ZoneID) %>%
  modelr::data_grid(corn = modelr::seq_range(corn, n = 101), 
                    maxdepth = modelr::seq_range(corn, n = 101), 
                    soilvorgvcarbon = modelr::seq_range(soilvorgvcarbon, n = 101), 
                    nitrogenvfertilizervuse = modelr::seq_range(nitrogenvfertilizervuse, n = 101), 
                    buffervcultivatedvcrops = modelr::seq_range(buffervcultivatedvcrops, n = 101), 
                    hu4vzoneid) %>%
  tidybayes::add_predicted_draws(fit_tn) %>%
  ggplot(aes(x = corn, y = tn)) +
  stat_lineribbon(aes(y = .prediction), 
                  .width = c(.99, .95, .8, .5), color = "#08519C") +
  geom_point(data = dplyr::filter(dt, hu4vzoneid %in% hu4_signif$ZoneID), 
                                  size = 2) +
  scale_fill_brewer() +
  facet_wrap(~ hu4vzoneid)

```

```{r residual_dotplot, echo=FALSE, eval=FALSE}
suppressMessages(library(tidybayes))
suppressMessages(library(brms))

lg <- lagosne_load("1.087.1")

fit <- readRDS("../data/mcmc/re/tn_corn")

dt        <- readRDS("../data/dt_scaled.rds") 
names(dt) <- gsub("_", "v", names(dt))

good_hu4s <- readRDS("../data/dt.rds") %>%
  group_by(hu4_zoneid) %>% 
  tally() %>%
  dplyr::filter(n >= 3)
dt <- dplyr::filter(dt, hu4vzoneid %in% good_hu4s$hu4_zoneid)

get_residuals <- function(model, threshold = 0.1){
  # model <- fit
  model$res_med <- dt %>%
    add_residual_draws(model) %>%
    group_by(lagoslakeid) %>%
    summarize(.residual_median = median(.residual)) %>%
    left_join(dplyr::select(lg$locus, lagoslakeid, nhd_lat, nhd_long),
              by = "lagoslakeid")

  model$res_med <- dt %>%
    add_fitted_draws(model) %>%
    group_by(lagoslakeid) %>%
    summarize(.value_median = median(.value)) %>%
    right_join(model$res_med, by = "lagoslakeid")

  model$res_test <- abs(
    median(model$res_med$.residual_median, na.rm = TRUE)) < threshold
  model
}
lg <- lagosne_load("1.087.1")
fit <- get_residuals(fit)

states_lg <- state_sf()[
  unlist(lapply(
    st_intersects(state_sf(), LAGOSNE::coordinatize(fit$res_med)),
    function(x) length(x) > 0)),]

ggplot() + 
  geom_sf(data = states_lg) +
  geom_sf(data = LAGOSNE::coordinatize(fit$res_med), 
          aes(color = .residual_median), size = 0.8) +
  coord_sf(datum = NA) +
  theme_void() + 
  scale_color_gradient2() +
  labs(color = "tn_corn \n model residuals")

```

```{r model_fitting, echo=FALSE, eval=FALSE}
library(brms)
library(tidybayes)
library(modelr)

brm_fit <- function(destfile, formula, data){
  fit <- brm(formula = formula, data = data, 
             family = gaussian())
  saveRDS(fit, destfile)
  return(fit)
}

# linear
model_form <- bf(tp ~ ag)
# unlink("../data/mcmc/fit0.rds")
fit0 <- get_if_not_exists(brm_fit, "../data/mcmc/fit0.rds", formula = model_form, 
                            data = dt)
print("fixed effect")
print(fit0$formula)

# linear re
model_form <- bf(tp ~ (1 + ag | hu4vzoneid))
# unlink("../data/mcmc/fit.rds")
fit <- get_if_not_exists(brm_fit, "../data/mcmc/fit.rds", formula = model_form, 
                         data = dt)
# plot(fit)
# print(fit$formula)

# linear re with focal predictor
model_form <- bf(tp ~ maxdepth + (1 + ag | hu4vzoneid))
# unlink("../data/mcmc/fit1.rds")
fit1 <- get_if_not_exists(brm_fit, "../data/mcmc/fit1.rds", formula = model_form, 
                         data = dt)
# plot(fit1)
# print(fit$formula)

model_form <- bf(tp ~ hu12vbaseflowvmean +
                   (1 + ag | hu4vzoneid))
# unlink("../data/mcmc/fit2.rds")
fit2 <- get_if_not_exists(brm_fit, "../data/mcmc/fit2.rds", formula = model_form,
                         data = dt)

# linear re with multivariate focal predictor
model_form <- bf(tp ~ maxdepth + iwslavratio + hu12vbaseflowvmean +
                   (1 + ag | hu4vzoneid))
# unlink("../data/mcmc/fit3.rds")
fit3 <- get_if_not_exists(brm_fit, "../data/mcmc/fit3.rds", formula = model_form, 
                         data = dt)

model_form <- bf(tp ~ maxdepth + iwslavratio + hu12vbaseflowvmean +
                   (1 + soybeans | hu4vzoneid))
# unlink("../data/mcmc/fit4.rds")
fit4 <- get_if_not_exists(brm_fit, "../data/mcmc/fit4.rds", formula = model_form, 
                         data = dt)

model_form <- bf(tp ~ maxdepth + iwslavratio + hu12vbaseflowvmean + soybeans)
# unlink("../data/mcmc/fit5.rds")
fit5 <- get_if_not_exists(brm_fit, "../data/mcmc/fit5.rds", formula = model_form, 
                         data = dt)

loo_compare(loo(fit4), loo(fit5))

quantile(ranef(fit4)$hu4vzoneid[,,2], c(0, 0.05, 0.5, 0.95, 1))
plot(fit4)

brms::bayes_R2(fit2)
brms::bayes_R2(fit3)
brms::bayes_R2(fit4)

range(dt$iwslavratio)

range((dt$iwslavratio - mean(dt$iwslavratio)) / sd(dt$iwslavratio))
hist((dt$iwslavratio - mean(dt$iwslavratio)) / sd(dt$iwslavratio))
test <- dt %>% mutate_if(function(col) is.numeric(col) & 
                    !all(col == .$lagoslakeid) &
                    !all(col == .$hu4vzoneid), scale)

# non-linear re
model_form <- bf(tp ~ bagvpct * ag, 
                 bagvpct ~ 1 + 
                   (1|hu4vzoneid), 
                 nl = TRUE)
fit1 <- get_if_not_exists(brm_fit, "../data/mcmc/fit1.rds", formula = model_form, 
                          data = dt)
# plot(fit1)
print("random effects")
print(fit1$formula)

# non-linear re w/ 2nd level coeffs
model_form2 <- bf(tp ~ bagvpct * ag, 
                 bagvpct ~ 1 + phosphorusvfertilizervuse + soybeans +
                   (1|hu4vzoneid), 
                 nl = TRUE)
fit_nle <- get_if_not_exists(brm_fit, "../data/mcmc/fit_nle.rds", formula = model_form2, 
                          data = dt)
print("re w/ 2nd level coeffs")
print(fit_nle$formula)

```

```{r model_diagnostics_basics, echo=FALSE, fig.height=2, eval=FALSE}
# get_variables(fit0)
plot(fit0, pars = "ag")
```

```{r model_diagnostics_advanced, fig.height=4, echo=FALSE, eval=FALSE}
# pull global distribution (lower, mean, upper)
# get_variables(fit0)
global_dist <- spread_draws(fit0, b_ag) %>%
    mutate(hu4vzoneid = "global") %>%
    group_by(hu4vzoneid) %>%
    do(tibble::as_tibble(t(quantile(.$b_ag, c(0.05, 0.5, 0.95)))))

# pull hu4 distribution (lower, mean, upper)
hu4_dist <- spread_draws(fit, r_hu4vzoneid[hu4vzoneid,term]) %>%
  dplyr::filter(term == "ag") %>%
  group_by(hu4vzoneid) %>%
  do(tibble::as_tibble(t(quantile(.$r_hu4vzoneid, c(0.05, 0.5, 0.95))))) %>%
  mutate(signif = case_when(`5%` > 0 ~ TRUE))

all_dist <- bind_rows(global_dist, hu4_dist)

ggplot(data = all_dist) +
  geom_pointrange(aes(x = hu4vzoneid, y = `50%`, 
                      ymin = `5%`, ymax = `95%`, color = signif)) +
  geom_hline(yintercept = 0) +
  coord_flip()

# STOP
hu4_signif <-  LAGOSNEgis::query_gis(
  "HU4", "ZoneID", dplyr::filter(hu4_dist, signif)$hu4vzoneid) %>% 
  st_cast(to = "MULTIPOLYGON")
```

```{r signif_map, echo=FALSE, eval=FALSE}
ggplot() + 
  geom_sf(data = hu4) +
  geom_sf(data = hu4_signif, color = "red")
```

```{r model_summary, echo=FALSE, fig.height=3, eval=FALSE}

# pull 2nd level covariate distributions
coef_signif <- bind_rows(
  spread_draws(fit_nle, b_bagvpct_soybeans) %>%
    do(tibble::as_tibble(t(quantile(.$b_bagvpct_soybeans, 
                                    c(0.05, 0.5, 0.95))))), 
  spread_draws(fit_nle, b_bagvpct_phosphorusvfertilizervuse) %>%
    do(tibble::as_tibble(t(quantile(.$b_bagvpct_phosphorusvfertilizervuse, 
                                    c(0.05, 0.5, 0.95)))))
  )
coef_signif$var <- c("soybeans", "pfertilizer")

ggplot(data = coef_signif) +
  geom_pointrange(aes(x = var, y = `50%`, ymin = `5%`, ymax = `95%`)) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  coord_flip()

```

```{r asdf, echo=FALSE, eval=FALSE}

# plot prediction lines
# hu4_signif$ZoneID
hu4_id               <- hu4_signif$ZoneID[2]
conditions           <- data.frame(hu4vzoneid = hu4_id)
rownames(conditions) <- hu4_id

me_zone <- marginal_effects(fit1, conditions = conditions, 
                            re_formula = NULL, method = "predict")

ran_ef <- as.data.frame(ranef(fit1)$hu4vzoneid) %>%
  mutate(zoneid = rownames(.)) %>%
  dplyr::filter(zoneid == hu4_id)

# plot(me_zone, ncol = 1)$agvpct # + 
  # geom_abline(intercept = 0, slope = ran_ef$Estimate.bagvpct_Intercept + 
  #               broom::tidy(fit)$estimate[3]) + 
  # ggtitle(hu4_id)

# ggplot() + 
#     geom_point(data = dplyr::filter(dt, 
#                                   hu4vzoneid == hu4_id), 
#              aes(x = agvpct, y = tp)) + 
#   geom_abline(intercept = 0, slope = ran_ef$Estimate.bagvpct_Intercept + 
#                 broom::tidy(fit)$estimate[3]) +
#   ylim(0, 450) + xlim(40, 90) + ggtitle(hu4_id)

dplyr::filter(dt, hu4vzoneid %in% hu4_signif$ZoneID) %>%
  modelr::data_grid(ag = seq_range(ag, n = 101), hu4vzoneid) %>%
  add_predicted_draws(fit) %>%
  ggplot(aes(x = ag, y = tp)) +
  stat_lineribbon(aes(y = .prediction), .width = c(.99, .95, .8, .5), color = "#08519C") +
  geom_point(data = dplyr::filter(dt, hu4vzoneid %in% hu4_signif$ZoneID), 
                                  size = 2) +
  scale_fill_brewer() +
  facet_wrap(~ hu4vzoneid)

```

```{r not_signif, echo=FALSE, eval=FALSE}
hu4_not_signif_n <- sample(unique(
  pull(
    dplyr::filter(dt, !(hu4vzoneid %in% hu4_signif$ZoneID)),
    hu4vzoneid)), 5)

dplyr::filter(dt, hu4vzoneid %in% hu4_not_signif_n) %>%
  modelr::data_grid(ag = seq_range(ag, n = 101), hu4vzoneid) %>%
  add_predicted_draws(fit) %>%
  ggplot(aes(x = ag, y = tp)) +
  stat_lineribbon(aes(y = .prediction), .width = c(.99, .95, .8, .5), color = "#08519C") +
  geom_point(data = dplyr::filter(dt, hu4vzoneid %in% hu4_not_signif_n), 
                                  size = 2) +
  scale_fill_brewer() +
  facet_wrap(~ hu4vzoneid)

# loo significance testing
loo  <- loo(fit)
loo0 <- loo(fit0)
loo1 <- loo(fit1)
loo2 <- loo(fit2)
loo_compare(loo0, loo1)
loo_compare(loo1, loo2)
```